### 2014-11-03 thesis/litreview/litreview.tex
A whole bunch of updates including source code listing in the appendix
Lots of changes

This is the method that Gabbe et al. used in their study to identify factors
that contributed to trauma patient length of stay and use these factors to
subsequently predict length of stay \citep{Gabbe2005}. The prediction problem
they proposed was essentially a binary classification problem, to which
logistic regression is well suited \citep{Tu1996}: will length of stay be
\textit{less than 5 days}, or \textit{equal to 5 days}? They used a
\textit{backward} stepwise selection algorithm, whereby all available variables
were initially included in the model and then progressively eliminated if their
likelihood ratio (LR) test $P$-value was $>0.15$. Features could only be
re-included if $P<0.05$. For length of stay
prediction, this left them with 9 features out of the 15 that they originally
included. To validate their model, they used a data set collected from an
independent source which had slightly different statistical characteristics for
the various features and outcomes. Their length of stay prediction model was
found to have 95\% sensitivity and 24\% specificity, with an area under the ROC
curve (AUC) of 0.62. This indicates that the model is not able to discriminate
length of stay cases of less than 5 days and equal to 5 days (i.e., the two
desired outcomes) very well, and only does
slightly better than simply random guessing (which has AUC 0.50).

In a similar study, Dinh et al. sought to predict whether or not trauma
patients could be predicted to stay in hospital \textit{less than 2 days} or
\textit{greater than 2 days} \citep{Dinh2013a}.
They also used a logistic regression model and
selected their features using a stepwise selection algorithm with the same
$P$-value threshold (0.05). Even though we cannot compare their result to that
of Gabbe et al. above because the models were tested on completely different
data sets, it is interesting to note that the model obtained by Dinh et al.
had an AUC of 0.83 after bootstrap cross-validation\footnote{A technique of
sampling a single data set with replacement: e.g. given a set of 100 records,
we can create a bootstrap sample that also contains 100 records which contains
records from the original set drawn at random with replacement. This means that
some records may appear in the bootstrap sample more than once and some may not
appear at all. In the work of Dinh et al., they performed this sampling 500
times (thereby creating 500 data sets) to evaluate the performance of their
model}. This indicates that the model had much better discrimination between
the two values for the outcome. Again, while we cannot directly compare, the
better performance could be due to the availability of more relevant features
to the outcome: Gabbe et al. included very few features specific to trauma
such as place of injury and mechanism of injury \citep{Gabbe2005} whereas Dinh
et al. also had variables for certain kinds of injuries (such as if it was
penetrating), specific body regions that were affected, and whether operations
were required. Given the nature of the length of stay, it would seem that Dinh
et al. derived a model that had included features with more explanatory power
on the outcome, thereby obtaining better results. This idea is consistent with
Buchman et al., who used logistic regression to predict the probability of a
patient in a surgical intensive care unit staying greater than a
week \citep{Buchman1994}.

While logistic regression is simple to use and its results straightforward
to interpret, the above approaches have described a coarse two-class prediction
problem, namely that of predicting the length of stay less than or greater than
a certain threshold. A number of \textit{linear} regression models have been
explored in the literature that examine the task of predicting length of stay
as a continuous quantity.

The key difference between linear and logistic regression is the nature of the
dependent variable: in linear regression it is treated as continuous (i.e., it
can take on any value), whereas in logistic regression it is a dichotomy of
only two values. The works discussed above only considered predicting the
length of stay as two possible outcomes, less than or greater than a particular
threshold of time. We have already stressed the importance of accurate length
of stay prediction for hospitals, and therefore it makes sense that attempting
to predict length of stay as a continous variable could give us more accurate
and useful insight and facilitate better resource allocation.

Yang et al. \citep{Yang2010} used a linear regression model as the baseline for
performance benchmarking when trying to predict the length of stay of burn
patients. Interestingly, unlike the logistic regression models discussed above,
they did not perform any feature selection when constructing their models and
found that this gave better performance against models constructed using
correlation-based feature selection\footnote{Abbreviated CFS, correlation-based
feature selection is a technique of choosing which independent variables to
include in a model based on the hypothesis that good variable subsets should
be highly correlated with the dependent variable but not with each
other \citep{Hall2000}.}. Unlike other studies into length of stay prediction,
they also looked at predicting the length of stay at various stages of the
clinical process, rather than simply at admission. By making predictions
at the admission, acute, and post-treatment stages, they were able to decrease
their predictions' mean absolute error (MAE) from 9.532 days at admission, to
6.314 days at post-treatment, where MAE is defined as
\begin{equation*}
MAE = \dfrac{1}{n}\sum_{i=1}^n |\mathrm{PredictedLOS}_i-\mathrm{ActualLOS}_i|
\end{equation*}
and PredictedLOS$_i$ is the predicted length of stay for the $i$-th instance,
ActualLOS$_i$ is the actual length of stay for the $i$-th instance, and $n$
is the total number of instances. Their approach was very much appropriate to
their medical domain (burn patients) as they tended to have longer lengths of
stay (which is more suitable for linear regression rather than logistic
regression) and the extended period of stay means that there will be enough
data from several phases of the clinical process for such patients. It is
difficult to apply such a phase method to our problem of trauma length of
stay prediction, however, as there is no standard clinical process due to
the varying nature of trauma causes and the lengths of stay are typically not
as long.

@@

\begin{table}[h]
\begin{tabular}{l|llll}
& Low & Medium & High & Exit \\
\hline
Low & 0.309 & 0.223 & 0.024 & 0.444 \\
Medium & 0.137 & 0.513 & 0.096 & 0.254 \\
High & 0.086 & 0.364 & 0.379 & 0.171 \\
Exit & 0 & 0 & 0 & 1 \\
\end{tabular}
\caption{}
\end{table}
From the matrix, we can say that the probability of transitioning from state
`Medium' to state `Low' is 0.137. Note that each row sums to 1, and that `Exit'
is an absorbing state: once entered, one can only re-enter the `Exit' state
with 100\% certainty. This represents patients leaving the ICU. How did they
obtain these probabilities? From their data set, they classified each patient
with an appropriate state (Low, Medium, High) corresponding to the severity of
their illness based on a particular mortality score (PRISM), and did this daily
over a few months; this gave them patient records that were simply strings of
L, M and H. To calculate the probability of transitioning from `Medium' to
`Low', they counted all occurrences of `Medium' to `Low' transitions and
divided this by the number of all transitions out of `Medium', and repeated
this for every transition. Using the theory of Markov processes, they are able
to calculate the average time taken before a person enters the `Exit' state,
which is precisely the length of stay. They found that the Markov model fitted
the data well and produced predictions that were very close to the observed
results in the test set \citep{Kapadia2000}, indicating that using such models
justifies further investigation.

@@

\subsection{Other Statistical Techniques}
A number of other techniques have been used in length of stay prediction, which
we will describe briefly.
Yau et al. \citep{Yau2003} proposed a \textit{mixture regression model} in
modelling
length of stay to overcome the problem of correlated observations in hospital
data, which allowed them to predict neonatal length of stay across heterogenous
sub-groups (short-stay and long-stay). The idea is that these two sub-groups
have different length of stay patterns, and the length of stay of any patient
is just the probability of them being in one of these two groups (as well as
taking into account the underlying distribution of the group itself). Applying
this method to a set of neonatal length of stay data, they were able to
identify which input factors contributed to which component (short or long
stay) and derive useful insight for hospitals.

@@

\section{Data Mining Methods}
Data mining is the analysis of (often large) observational data sets to find
unsuspected relationships and to summarise the data in novel ways that are both
understandable and useful to the data owner \citep{Hand2001}. Compared to
traditional statistical, it is a relatively recent field that emerged in the
1990s, and has
given researchers a variety of new tools to analyse and understand biomedical
datasets \citep{Bellazzi2008,Yoo2012}. Bellazzi \& Zupan \citep{Bellazzi2008}
have written a
comprehensive review paper of data mining techniques in clinical prediction
settings, and we refer the reader to their article for a more detailed
treatment of the techniques and their applications. Here we will describe the
major methods in the context of their application to predicting hospital length
of stay.

@@

Yang et al. investigated the performance of linear regression, M5 (another
tree-based classification method), and support vector machines in predicting
length of stay for burn patients. The difference between their tree method,
M5, with that of Harper (CART), is that the nodes contain linear regression
functions instead of the usual discrete (two-class) outputs that are used to
successively split the data set. Additionally, they also use support vector
machines (SVMs), perhaps the most powerful classification algorithm in terms of
predictive accuracy \citep{Bellazzi2008}. SVMs work on the principle that when
the data is represented in a high enough dimensional space, a linear hyperplane
can be found where the margin between two different records is as large as
possible. In Figure \ref{fig:svm-example} from Yoo et al. \citep{Yoo2012},
hyperplane 2 is the better plane because its margin is as large as possible. 
\begin{figure}[h]
\label{fig:svm-example}
\centering
\includegraphics[scale=1.05]{images/a2-svm}
\caption{An example of finding the largest margin hyperplane for a two-class
support vector machine classifier.}
\end{figure}

>>>>

This is the method that Gabbe et al. used in their study to identify factors
that contributed to trauma patient length of stay and use these factors to
subsequently predict length of stay \citep{Gabbe2005}. The prediction problem
they proposed was essentially a binary classification problem, to which
logistic regression is well suited \citep{Tu1996}: will length of stay be
\textit{less than 5 days}, or \textit{equal to 5 days}? They used a
\textit{backward} stepwise selection algorithm, whereby all available variables
were initially included in the model and then progressively eliminated if their
likelihood ratio (LR) test $P$-value was $>0.15$. Features could only be
re-included if $P<0.05$. For length of stay
prediction, this left them with 9 features out of the 15 that they originally
included. To validate their model, they used a data set collected from an
independent source which had slightly different statistical characteristics for
the various features and outcomes. Their length of stay prediction model was
found to have 95\% sensitivity and 24\% specificity, with an area under the ROC
curve (AUC) of 0.62. This indicates that the model is not able to accurately
discriminate
length of stay cases of less than 5 days and equal to 5 days (i.e., the two
desired outcomes) very well, performing only
slightly better than simply random guessing (which has AUC 0.50).

In a similar study, Dinh et al. sought to predict whether or not trauma
patients could be predicted to stay in hospital for \textit{less than 2 days}
or \textit{greater than 2 days} \citep{Dinh2013a}.
They also used a logistic regression model and
selected their features using a stepwise selection algorithm with the same
$P$-value threshold (0.05). Even though we cannot directly compare their result
to that
of Gabbe et al. above, because the models were tested on completely different
data sets, it is interesting to note that the model obtained by Dinh et al.
had an AUC of 0.83 after bootstrap cross-validation\footnote{A technique of
sampling a single data set with replacement: e.g. given a set of 100 records,
we can create a bootstrap sample that also contains 100 records which contains
records from the original set drawn at random with replacement. This means that
some records may appear in the bootstrap sample more than once and some may not
appear at all. In the work of Dinh et al., they performed this sampling 500
times (thereby creating 500 data sets) to evaluate the performance of their
model}. This indicates that the model had much better discrimination between
the two values for the outcome. Again, while we cannot directly compare, the
better performance could be due to the presence of features in the data more
relevant to the outcome: the data of Gabbe et al. included very few features
specific to trauma \citep{Gabbe2005} whereas Dinh
et al. had variables indicating the kind of injury (such as penetrating trauma,
the bodily regions that were affected, and whether operations
were required). Given the nature of the length of stay it would seem that Dinh
et al. derived a model that had included features with more explanatory power
on the outcome, thereby obtaining better results. This idea is consistent with
Buchman et al., who used logistic regression to predict the probability of a
patient in a surgical intensive care unit staying greater than one
week \citep{Buchman1994}.

While logistic regression is simple to use and its results straightforward
to interpret, the above approaches describe a coarse two-class prediction
problem, namely that of predicting length of stay dichotomised by an
arbitrarily chosen threshold.
A number of \textit{linear} regression models have been
explored in the literature that examine the problem of predicting length of stay
as a continuous quantity.

The key difference between linear and logistic regression is the nature of the
dependent variable: in linear regression it is treated as continuous (i.e., it
can take on any value), whereas in logistic regression it is a dichotomy of
only two values. The works discussed above only considered predicting the
length of stay as two possible outcomes, less than or greater than a particular
threshold of time. We have already stressed the importance of accurate length
of stay prediction for hospitals, and therefore it makes sense that attempting
to predict length of stay as a continous variable could give us more accurate
and useful insights, and facilitate better resource allocation.

Yang et al. \citep{Yang2010} used a linear regression model as the baseline for
performance benchmarking when trying to predict the length of stay of burn
patients. Interestingly, unlike the logistic regression models discussed above,
they did not perform any feature selection when constructing their models and
found that this gave better performance compared with models constructed using
correlation-based feature selection\footnote{Abbreviated CFS, correlation-based
feature selection is a technique of choosing which independent variables to
include in a model based on the hypothesis that good variable subsets should
be highly correlated with the dependent variable but not with each
other \citep{Hall2000}.}. Unlike other studies into length of stay prediction,
they also looked at predicting the length of stay at various stages of the
clinical process, rather than simply at admission. By making predictions
at the admission, acute, and post-treatment stages, they were able to decrease
the mean absolute error (MAE) of their predictions from 9.532 days at admission,
to 6.314 days at post-treatment, where MAE is defined as
\begin{equation*}
MAE = \dfrac{1}{n}\sum_{i=1}^n |\mathrm{PredictedLOS}_i-\mathrm{ActualLOS}_i|
\end{equation*}
and PredictedLOS$_i$ is the predicted length of stay for the $i$-th instance,
ActualLOS$_i$ is the actual length of stay for the $i$-th instance, and $n$
is the total number of instances. Their approach was very much appropriate to
their medical domain (burn patients) as they tended to have longer lengths of
stay (which is more suitable for linear regression rather than logistic
regression) and the extended period of stay means that there will be enough
data from several phases of the clinical process for such patients. It is
difficult to apply such a phase method to our problem of trauma length of
stay prediction, however, as there is no standard clinical process due to
the varying nature of trauma causes and the lengths of stay are typically not
as long.

@@

\begin{table}[h]
\begin{tabular}{l|llll}
& Low & Medium & High & Exit \\
\hline
Low & 0.309 & 0.223 & 0.024 & 0.444 \\
Medium & 0.137 & 0.513 & 0.096 & 0.254 \\
High & 0.086 & 0.364 & 0.379 & 0.171 \\
Exit & 0 & 0 & 0 & 1 \\
\end{tabular}
\caption{A transition probability matrix from Kapadia et al. for the
movement of patients through a pediatric intensive care unit.}
\end{table}
From the matrix, we can say that the probability of transitioning from state
`Medium' to state `Low' is 0.137. Note that each row sums to 1, and that `Exit'
is an absorbing state: once entered, one can only re-enter the `Exit' state
with 100\% certainty. This represents patients leaving the ICU. How did they
obtain these probabilities? From their data set, they classified each patient
with an appropriate state (Low, Medium, High) corresponding to the severity of
their illness based on a particular mortality score (PRISM), and did this daily
over a few months; this gave them patient records that were simply strings of
L, M and H. To calculate the probability of transitioning from `Medium' to
`Low', they counted all occurrences of `Medium' to `Low' transitions and
divided this by the number of all transitions out of `Medium', and repeated
this for every transition. Using the theory of Markov processes, they are able
to calculate the average time taken before a person enters the `Exit' state,
which is precisely the length of stay. They found that the Markov model fitted
the data well and produced predictions that were very close to the observed
results in the test set \citep{Kapadia2000}, indicating that using such models
justifies further investigation.

@@

\subsection{Other Statistical Techniques}
A number of other techniques have been used in length of stay prediction, which
we will describe briefly.
Yau et al. \citep{Yau2003} proposed a \textit{mixture regression model} in
modelling
length of stay to overcome the problem of correlated observations in hospital
data, which allowed them to predict neonatal length of stay across heterogenous
sub-groups (short-stay and long-stay). Their idea is that these two sub-groups
have different length of stay patterns, and the length of stay of any patient
is just the probability of them being in one of these two groups (as well as
taking into account the underlying distribution of the group itself). Applying
this method to a set of neonatal length of stay data, they were able to
identify which input factors contributed to which component (short or long
stay) and thereby derive a useful insight for hospitals.

@@

\section{Data Mining Methods}
Data mining is the analysis of (often large) observational data sets to find
unsuspected relationships and to summarise the data in novel ways that are both
understandable and useful to the data owner \citep{Hand2001}. Compared to
traditional statistical techniques, it is a relatively recent field, having
emerged in the 1990s, and has
given researchers a variety of new tools to analyse and understand biomedical
datasets \citep{Bellazzi2008,Yoo2012}. Bellazzi \& Zupan \citep{Bellazzi2008}
have written a
comprehensive review paper of data mining techniques in clinical prediction
settings, and we refer the reader to their article for a more detailed
treatment of the techniques and their applications. Here we will describe the
major methods in the context of their applicability to the problem of
predicting length of hospital stay.

@@

Yang et al. investigated the performance of linear regression, M5 (another
tree-based classification method), and support vector machines in predicting
length of stay for burn patients. The difference between their tree method,
M5, with that of Harper (CART), is that the nodes contain linear regression
functions instead of the usual discrete (two-class) outputs that are used to
successively split the data set. Additionally, they also use support vector
machines (SVMs), perhaps the most powerful classification algorithm in terms of
predictive accuracy \citep{Bellazzi2008}. SVMs work on the principle that when
the data is represented in a high enough dimensional space, a linear hyperplane
can be found where the margin between two different records is as large as
possible. In Figure \ref{fig:svm-example} from Yoo et al. \citep{Yoo2012},
hyperplane 2 is the better plane because its margin is as large as possible. 
\begin{figure}[h]
\centering
\includegraphics[scale=1.05]{images/a2-svm}
\caption{An example of finding the largest margin hyperplane for a two-class
support vector machine classifier.}
\label{fig:svm-example}
\end{figure}

### 2014-11-03 thesis/method/classification.tex
Add SVM diagram
Add MLP diagram
Fix strange scaling of SVM, MLP and feature selection diagrams
Lots of changes

To find the attribute to split on, we assume we are splitting on a particular
attribute, and calculate the information at the node given the split. We do
this for each attribute, and compute the \textit{information gain} from
splitting on that attribute:
\begin{equation*}
\mathrm{gain}(feature) = \mathrm{info}(\text{child node}) - \mathrm{info}(\text{parent node})
\end{equation*}
The feature that is chosen is therefore the one with the highest information
gain.

@@

Given a set of $n$-feature vectors, we can represent each vector as a point in
$n$-dimensional space, where each feature is a dimension. The SVM will try to
find a \textit{hyperplane} that separates the two classes. Usually there are
many such hyperplanes; the distinguishing aspect of an SVM is that it will
find the hyperplane that maximises the distance between the two nearest points
in each class. This distance is called the \textit{margin}, and the intuition
is that maximising it will give us the greatest possible separation between the
classes -- the classifier comes no closer to either class than it needs to.
This gives the SVM the best generalising ability to classify new examples.

@@

\begin{figure}[h]
\caption{An example of a 2-dimensional SVM.}
\label{fig:svm}
\centering
%\includegraphics[scale=0.95]{images/a2-nn}
\end{figure}

@@

@@

Like the Na\"{i}ve Bayes classifier, $k$-NN allows all features to be taken
into account equally when classifying a new example. However, since feature
values often span different ranges and units of measurement, a particularly
large range of values for a feature would contribute much more to the distance
than another feature that spans a lower set of values.
To avoid this, it is important to \textit{normalise} the
values of each numeric feature before using $k$-NN: for each feature $i$, we
calculate the normalised value of the feature for each training example. This
is done using the following relationship:
\begin{equation*}
a_i = s\left(\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right) + t
\end{equation*}
where $a_i$ is the normalised value, $v_i$ is the actual value in the data
set, max $v_i$ and min $v_i$ are taken over all training examples, $s$ is a
scaling factor $t$ indicates how much to translate the range by. If $s=1$
and $t=0$, then the range of normalised values will be in $[0,1]$.

\subsection{Existing and Proposed Extensions}
There have been many extensions proposed to the $k$-NN algorithm described
above. These usually seek to find a more suitable distance metric for the
problem domain at hand, or to reduce the storage requirements of the original
$k$-NN algorithm. We will focus on two extensions to the Euclidean distance:
K* and our proposed distance function, Ranked Distance.

@@

We use K* as a comparison to the basic NN algorithm and also as a comparison
to $k$-NN with our distance function.

\subsubsection{Proposed Approach: Ranked Distance}
\label{sec:rankeddist}
\paragraph{Intuition}
As we mentioned earlier, all features are considered equally in a $k$-NN
classifier, which implies that all features are equally important in predicting
the class. In practice, however, this is not always the case: to predict the
LOS of a trauma patient, the severity of their injury should affect the final
LOS more than whether or not they can speak English (which could be just two
of many features that are recorded about a patient). Our contribution is thus
a modified distance function to take the relative importance of attributes into
account.

@@

\paragraph{Assignment of weights}
The weights $w_i$ can be tuned to match the particular problem that is being
investigated. For our LOS classification problem, we will consider two ways
of weighting features:
\begin{enumerate}
\item The weight of each feature is the magnitude of the correlation
coefficient
between it and the class, meaning that features that are more highly correlated
with the class will contribute more to the distance. The correlation
coefficient $r_i$ for feature $i$ with the class $c$ can be computed by:
\begin{equation*}
r_i = \dfrac{\sum_{i=1}^m (x_i-\bar{x})(c-\bar{c})}{\sqrt{(\sum_{i=1}^m x_i-\bar{x})(\sum_{i=1}^m c-\bar{c})}}
\end{equation*}
where $m$ is the number of training examples, $x_i$ is the value of feature
$i$, and $\bar{x}$ and $\bar{c}$ are the arithmetic averages of the feature
values and the class values respectively. 
\item We compute the correlation coefficients as described in the point above.
This gives us a ranking of the importance of each feature to the prediction of
the class, with 1 being the highest rank (indicating the most importance) and
$n$ being the lowest rank (indicating the least importance).
Instead of using the values of the correlation directly to weight
the distance contribution of each feature, we specify a function:
\begin{equation*}
f : \mathrm{Rank} \rightarrow \mathbb{R}, \mathrm{Rank} \in \{1,2,\ldots,n\}
\end{equation*}
that describes
how the weights of the features vary with the rank. This function should
inituitively be non-increasing and should produce lower values when the rank
number is greater, meaning that less important features have a lower weight
in the distance calculation.
Note that choosing $f(\mathrm{Rank}) = k$ for any non-zero constant $k$
results in all features contributing equally to the distance calculation.

\noindent We will use $f(\mathrm{Rank}) = \frac{1}{\mathrm{Rank}^c}$, where
$c$ varies from 0 to 1.
\end{enumerate}

\begin{figure}[h]
\caption{An MLP with one hidden layer.}
\label{fig:mlp}
\centering
%\includegraphics[scale=0.95]{images/a2-nn}
\end{figure}

\section{Summary}
In this chapter we have outlined all of the learning algorithms that we will
be evaluating with respect to our classification problem, from the simple
Zero-R and 1R to Na\"{i}ve Bayes, decision trees, logistic regression,
$k$-nearest neighbour (and extensions K* and Ranked Distance), to the more
sophisticated support vector machines and multi-layer perceptrons. The reason
for our inclusion of so many algorithms is two-fold. Firstly, many of
these algorithms have not been extensively applied in various medical settings
to the problem of LOS prediction. Using a wide range of algorithms from various
paradigms will enable us to see what algorithms perform better than others.
Secondly, evaluating several classifiers will allow us to make comparisons
between different feature selection methods that are independent of the
classifier and give us more insight into the nature of the data set and of the
problem.

>>>>

To find the attribute to split on, we assume we are splitting on a particular
attribute, and calculate the information at the node given the split. We do
this for each attribute, and compute the \textit{information gain} from
splitting on that attribute:
\begin{equation*}
\mathrm{gain}(feature) = \mathrm{info}(\text{split}) - \mathrm{info}(\text{parent})
\end{equation*}
The feature that is chosen is therefore the one with the highest information
gain.

@@

Given a set of $n$-feature vectors, we can represent each vector as a point in
$n$-dimensional space, where each feature is a dimension. The SVM will try to
find a \textit{hyperplane} that separates the two classes. Usually there are
many such hyperplanes; the distinguishing aspect of an SVM is that it will
find the hyperplane that maximises the distance between the two nearest points
in each class. This distance is called the \textit{margin}, and the intuition
is that maximising it will give us the greatest possible separation between the
classes -- the classifier comes no closer to either class than it needs to.
This gives the SVM the best generalising ability to classify new examples.
\begin{figure}[h]
\includegraphics{images/method/svm-example.pdf}
\caption{An example of a 2-dimensional SVM. The solid line indicates the
maximum margin hyperplane.}
\label{fig:svm}
\end{figure}

@@

@@

\begin{figure}[h]
\centering
\includegraphics{images/method/mlp.pdf}
\caption{An MLP with one hidden layer. There are $n$ input nodes, $k$ nodes
in the hidden layer, and one output node.}
\label{fig:mlp}
\end{figure}

@@

Like the Na\"{i}ve Bayes classifier, $k$-NN allows all features to be taken
into account equally when classifying a new example. However, since feature
values often span different ranges and units of measurement, a particularly
large range of values for a feature would contribute much more to the distance
than another feature that spans a lower set of values.
To avoid this, it is important to \textit{normalise} the
values of each numeric feature before using $k$-NN: for each feature $i$, we
calculate the normalised value of the feature for each training example. This
is done using the following relationship:
\begin{equation*}
a_i = s\left(\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right) + t
\end{equation*}
where $a_i$ is the normalised value, $v_i$ is the actual value in the data
set, max $v_i$ and min $v_i$ are taken over all training examples, $s$ is a
scaling factor and $t$ indicates how much to translate the range by. If $s=1$
and $t=0$, then the range of normalised values will be in $[0,1]$.

\subsection{Existing and Proposed Extensions}
There have been many extensions proposed to the $k$-NN algorithm described
above. These usually seek to find a more suitable distance metric for the
problem domain at hand, or to reduce the storage requirements of the original
$k$-NN algorithm. We will focus on two extensions to the Euclidean distance:
K* and our proposed algorithm, Ranked Distance.

@@

We use K* as a comparison to the basic NN algorithm and also as a comparison
to our proposed Ranked Distance Nearest Neighbours algorithm.

\subsubsection{Proposed Approach: Ranked Distance Nearest Neighour}
\label{sec:rankeddist}
\paragraph{Intuition}
As we mentioned earlier, all features are considered equally in a $k$-NN
classifier, which implies that all features are equally important in predicting
the class. In practice, however, this is not always the case: to predict the
LOS of a trauma patient, the severity of their injury should affect the final
LOS more than whether or not they can speak English (which could be just two
of many features that are recorded about a patient). Our contribution is a
new NN algorithm that uses a modified distance function which takes into
account the relative importance of all features called Ranked Distance
Nearest Neighbour (RD).

@@

\paragraph{Assignment of weights}
The weights $w_i$ can be tuned to match the particular problem that is being
investigated. For our LOS classification problem, we will consider one way
of weighting features.
First, we compute the correlation coefficients using:
\begin{equation*}
r_i = \dfrac{\sum_{i=1}^m (x_i-\bar{x})(c-\bar{c})}{\sqrt{(\sum_{i=1}^m x_i-\bar{x})(\sum_{i=1}^m c-\bar{c})}}
\end{equation*}
where $m$ is the number of training examples, $x_i$ is the value of feature
$i$, and $\bar{x}$ and $\bar{c}$ are the arithmetic averages of the feature
values and the class values respectively. 
This gives us a ranking of the importance of each feature to the prediction of
the class, with 1 being the highest rank (indicating the most importance) and
$n$ being the lowest rank (indicating the least importance).
Instead of using the values of the correlation directly to weight
the distance contribution of each feature, we specify a function:
\begin{equation*}
f : \mathrm{Rank} \rightarrow \mathbb{R}, \mathrm{Rank} \in \{1,2,\ldots,n\}
\end{equation*}
that describes
how the weights of the features vary with the rank. This function should
inituitively be non-increasing and should produce lower values when the rank
number is greater, meaning that less important features have a lower weight
in the distance calculation.
Note that choosing $f(\mathrm{Rank}) = k$ for any non-zero constant $k$
results in all features contributing equally to the distance calculation.
We will use $f(\mathrm{Rank}) = \frac{1}{\mathrm{Rank}^c}$, where
$c$ varies from 0 to 1.

\section{Summary}
In this chapter we have outlined all of the learning algorithms that we will
be evaluating with respect to our classification problem, from the simple
Zero-R and 1R to Na\"{i}ve Bayes, decision trees, logistic regression,
$k$-nearest neighbour (and K*), to the more
sophisticated support vector machines and multi-layer perceptrons. We also
described our proposed new algorithm, Ranked Distance Nearest Neighbours.
The reason
for our inclusion of so many algorithms is two-fold. Firstly, many of
these algorithms have not been extensively applied in various medical settings
to the problem of LOS prediction. Using a wide range of algorithms from various
paradigms will enable us to see what algorithms perform better than others.
Secondly, evaluating several classifiers will allow us to make comparisons
between different feature selection methods that are independent of the
classifier and give us more insight into the nature of the data set and of the
problem.
